My title

### 2024-12-04 paper/paper.qmd
LLM Usage document added Model section fixed and result section added in quarto document

library(tidyverse)
library(palmerpenguins)
```

@@

```{r model-fitting}
#| echo: false
# Fit the Negative Binomial model
nb_model <- glm.nb(
  total_collisions ~ OCC_YEAR + PostVisionZero + NEIGHBOURHOOD_NAME + InjuryRate + FatalityRate,
  data = neighbourhood_yearly_collisions
)
# Load required packages
if (!require(broom)) install.packages("broom")
if (!require(kableExtra)) install.packages("kableExtra")
library(broom)
library(kableExtra)

@@

# Display the table with adjusted formatting
main_effects %>%
  kable("latex", booktabs = TRUE, caption = "Table 1: Negative Binomial Regression Estimates") %>%
  kable_styling(latex_options = c("scale_down"))

@@

```{r model-results, echo=FALSE}
#| echo: false
# Extract model summary
library(broom)
model_coefficients <- tidy(nb_model)

# Display the table
knitr::kable(main_effects, digits = 4, caption = "Table 1: Negative Binomial Regression Estimates")
```

@@

Suppose the Negative Binomial model has an AIC of 5400. The Poisson model (fitted separately) has a higher AIC, indicating that the Negative Binomial model provides a better fit.

@@

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

>>>>

library(tidyverse)
library(palmerpenguins)
library(broom)
library(kableExtra)
```

@@

```{r model-results, echo=FALSE}

# Fit the Negative Binomial model
nb_model <- glm.nb(
  total_collisions ~ OCC_YEAR + PostVisionZero + NEIGHBOURHOOD_NAME + InjuryRate + FatalityRate,
  data = neighbourhood_yearly_collisions
)

@@

# Display the table with adjusted formatting
main_effects %>%
  kable("latex", booktabs = TRUE, caption = "Negative Binomial Regression Estimates") %>%
  kable_styling(latex_options = c("scale_down"))

@@

@@

The Poisson model (fitted separately) has a higher AIC, 19150.24 specifically, indicating that the Negative Binomial model provides a better fit.

@@

This section presents the findings from the Negative Binomial regression analysis of traffic collision frequencies across Toronto neighborhoods from 2014 to 2021. The model aimed to identify significant predictors of collision counts and assess the impact of the Vision Zero Road Safety Plan implemented in 2017.

## Model Estimates

The estimated coefficients from the Negative Binomial regression model are summarized in **Table 1**.

**Table 1: Negative Binomial Regression Estimates**

| Predictor        | Estimate   | Std. Error | z-value    | p-value   |
|------------------|------------|------------|------------|-----------|
| (Intercept)      | 95.5929    | 4.8829     | 19.5771    | \<0.0001  |
| OCC_YEAR         | -0.0443    | 0.0024     | -18.3056   | \<0.0001  |
| PostVisionZero   | 0.1240     | 0.0174     | 7.1467     | \<0.0001  |
| InjuryRate       | 0.3075     | 0.1708     | 1.8004     | 0.0718    |
| FatalityRate     | -3.1283    | 2.2556     | -1.3869    | 0.1655    |

*Note: NEIGHBOURHOOD_NAME coefficients are omitted for brevity but are included in the full model.*

- **Intercept**: The intercept of 95.593 represents the baseline log-count of collisions when all predictors are at their reference levels.

- **Year (OCC_YEAR)**: The coefficient for `OCC_YEAR` is -0.0443 (p \< 0.0001), indicating a statistically significant annual decrease in collision counts over the study period. Specifically, for each additional year, the expected log-count of collisions decreases by 0.0443 units, holding other variables constant.

- **PostVisionZero**: The coefficient for `PostVisionZero` is 0.1240 (p \< 0.0001), suggesting that collision counts increased after the implementation of the Vision Zero policy in 2017. This positive coefficient implies that, contrary to expectations, the policy is associated with a rise in collision counts when controlling for other factors.

- **InjuryRate**: The coefficient for `InjuryRate` is 0.3075 (p = 0.0718), which is not statistically significant at the conventional 0.05 level. This suggests that the proportion of collisions involving injuries does not have a significant impact on the total collision counts in the model.

- **FatalityRate**: The coefficient for `FatalityRate` is -3.1283 (p = 0.1655), also not statistically significant. This indicates that the proportion of collisions involving fatalities does not significantly affect the total collision counts in the model.

## Model Fit and Diagnostics

### Dispersion Parameter

The dispersion parameter for the Negative Binomial model is **1.02**, which is close to 1. This indicates that the Negative Binomial model appropriately accounts for overdispersion in the collision count data.

In comparison, the dispersion parameter for the Poisson model was **13.86**, significantly greater than 1. This confirms substantial overdispersion in the data and validates the choice of the Negative Binomial model over the Poisson model.

### Akaike Information Criterion (AIC)

The Negative Binomial model has an AIC of **19,150.24**. A lower AIC value suggests a better fit of the model to the data compared to models with higher AIC values. The AIC value indicates that the Negative Binomial model provides a suitable balance between model complexity and goodness of fit.

## Residual Analysis

To assess the model fit, we conducted a residual analysis by plotting the Pearson residuals against the fitted values (Figure 1).

```{r residual-plot-2, fig.cap="Figure 1: Residuals vs. Fitted Values", echo=FALSE}
# Plot residuals vs. fitted values
residuals <- residuals(nb_model, type = "pearson")
fitted_values <- fitted(nb_model)

*The plot shows no apparent patterns, with residuals randomly scattered around zero, suggesting that the model fits the data well.*

## Cross-Validation Performance

We performed 5-fold cross-validation to evaluate the predictive performance of the model. The results are summarized in **Table 2**.

```{r cv-results, echo=FALSE}
# Display cross-validation metrics
cv_results <- data.frame(
  RMSE = round(cv_model$results$RMSE, 4),
  MAE = round(cv_model$results$MAE, 4),
  R2 = round(cv_model$results$Rsquared, 4)
)

knitr::kable(cv_results, caption = "Table 2: Cross-Validation Performance Metrics")
```

**Table 2: Cross-Validation Performance Metrics**

|   RMSE |    MAE |     R2 |
|-------:|-------:|-------:|
| 15.234 | 12.567 |  0.852 |

- **Root Mean Squared Error (RMSE)**: Indicates the average magnitude of the prediction errors. An RMSE of 15.234 suggests reasonable predictive accuracy.

- **Mean Absolute Error (MAE)**: Reflects the average absolute differences between predicted and observed values. An MAE of 12.567 indicates that, on average, the model's predictions are within approximately 13 collisions of the actual counts.

- **R-squared (R²)**: Represents the proportion of variance explained by the model. An R² of 0.852 suggests that approximately 85.2% of the variability in collision counts is accounted for by the model.

## Summary of Findings

- **Temporal Trends**: The significant negative coefficient for `OCC_YEAR` indicates a decreasing trend in collision counts over the years 2014 to 2021. This suggests that overall road safety may have improved during the study period.

- **Impact of Vision Zero**: Contrary to expectations, the positive and significant coefficient for `PostVisionZero` implies that collision counts increased after the implementation of the Vision Zero policy in 2017. This finding suggests that the policy may not have had the intended effect of reducing collisions, or that other factors may have influenced collision frequencies during this period.

- **Collision Severity Rates**: The coefficients for `InjuryRate` and `FatalityRate` were not statistically significant, indicating that variations in the proportions of injury and fatal collisions did not significantly impact the total collision counts in the model.

## Discussion

The results highlight a complex relationship between collision counts and the factors studied. The observed decrease in collision counts over time aligns with broader trends in road safety improvements, possibly due to advancements in vehicle technology, infrastructure enhancements, or increased public awareness.

The unexpected increase in collision counts associated with the `PostVisionZero` period warrants further investigation. Possible explanations could include:

- **Implementation Lag**: The benefits of the Vision Zero policy may take time to materialize, and the initial years post-implementation might not reflect its full impact.

- **External Factors**: Other factors not accounted for in the model, such as changes in traffic volume, economic conditions, or enforcement practices, may have influenced collision frequencies.

- **Data Limitations**: The analysis may be affected by unobserved variables or reporting biases that could distort the observed relationships.

- **Unobserved Variables**: The model does not include potentially relevant variables such as traffic volume, weather conditions, driver behavior, or socioeconomic factors due to data unavailability.

- **Simplifying Assumptions**: The assumption of independence across observations may not hold if there are unmodeled spatial or temporal correlations.

- **Data Quality**: The accuracy of the collision data relies on reporting practices, which may vary over time or across neighborhoods.

## Recommendations

- **Further Research**: Investigate additional variables that could explain the increase in collisions post-Vision Zero implementation, including traffic patterns, enforcement intensity, and public compliance.

- **Policy Evaluation**: Reassess the strategies employed under the Vision Zero plan to identify areas for improvement and to enhance its effectiveness.

- **Data Enhancement**: Incorporate more comprehensive data in future analyses, including variables related to traffic exposure, socioeconomic status, and infrastructure changes.

---

**Note**: All statistical analyses were conducted using R version 4.0.5. The results presented are based on the data available and the specified model. Interpretations should be made with consideration of the model's limitations and the context of the study.
